---
title: "Project 1: Reading JSON Files in R"
author: 
- "Min He"
date: "June 9, 2020 (updated `r Sys.Date()`)"
output:
  html_document: 
    number_sections: yes
    toc: yes
    toc_depth: 3
    code_folding: show
    toc_float:
      collapsed: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE)
library(tidyverse)
library(jsonlite)
library(dplyr)

require(knitr)
library(ggplot2)

library(DT)
```
# JSON Description
JSON (JavaScript Object Notation) is a lightweight data-interchange format. It is easy for humans to read and write. It is easy for machines to parse and generate. JSON is common data formats used to receive data from a web server. JSON is built on two structures: 
1. A collection of name/value pairs.
2. An ordered list of values. 


# Pros/Cons of Packages

We chose the `jsonlite` package to read in our data [@jsonlite]. The `rjson` package was also considered, but inital attempts performed slowly and needed extra code to implement the data into a dataframe [@rjson]. The `jsonlite` package converts read data directly into a dataframes by default, while `rjson` will convert data into a list by default. Given our need to work with large data sets and to create dataframes, `jsonlite` performs better for our needs.


# Data

This data we are using is the NHL records, an jsonlite API is used to retrieve all the team data (including the details about team New Jersey Devils).

## Read the data
```{r echo=TRUE}
# Read in data from JSON source

root_url <- 'https://records.nhl.com/site/api'
json_file <- (paste0(root_url,'/franchise'))
source_data <- fromJSON(json_file)
franchise_df <- as.data.frame(source_data$data) #Convert to data frame

json_file <- (paste0(root_url,'/franchise-team-totals'))
source_data <- fromJSON(json_file)
franchise_team_total <- as.data.frame(source_data$data) #Convert to data frame
franchise_team_total$gameTypeId <- as.factor(franchise_team_total$gameTypeId)


franchise_id=23
json_file <- (paste0(root_url,'/franchise-season-records?cayenneExp=franchiseId=',franchise_id))
source_data <- fromJSON(json_file)
franchise_team_season <- as.data.frame(source_data$data) #Convert to data frame


json_file <- (paste0(root_url,'/franchise-goalie-records?cayenneExp=franchiseId=',franchise_id))
source_data <- fromJSON(json_file)
franchise_team_goalie <- as.data.frame(source_data$data) #Convert to data frame

json_file <- (paste0(root_url,'/franchise-skater-records?cayenneExp=franchiseId=',franchise_id))
source_data <- fromJSON(json_file)
franchise_team_skater <- as.data.frame(source_data$data) #Convert to data frame

```
## Create function (Creates Labels for Variable Names)

A basic exploratory data analysis.
+ The scatterplot between losses and home losses shows a very good linear relationship
+ The linear relationship between losses and home losses still hold within each game type.

```{r echo=TRUE}

ggplot(franchise_team_total,aes(losses,homeLosses)) +
  geom_point()


ggplot(franchise_team_total,aes(losses,homeLosses)) +
  geom_point() +
  facet_grid(cols=vars(gameTypeId),scales = "free",labeller = label_both)
```

**Take the** *franchiseId=23* **an an example, the scatterplot between most wins in one season and total wins for each team member shows some linear relationship and this relationship is better when the outlier was excluded.**
```{r}
plot(franchise_team_goalie$mostWinsOneSeason, franchise_team_goalie$wins, main='Relationshio between most wins in one season\n v.s. wins for each member')
plot(franchise_team_goalie$mostWinsOneSeason[-1], franchise_team_goalie$wins[-1], main='Relationshio between most wins in one season\n v.s. wins for each member (exclude outlier)')
```

## Create Function (Generates Contingency Table)
**Show contingency table with columns of team and game type (summarise the total of wins)**
```{r, echo=TRUE}
#Function creates Contingency Table
contingencyTable<-franchise_team_total %>% group_by_at(c('teamName','gameTypeId')) %>% 
    summarise(wins = sum(wins)) %>% 
    spread(key = 'gameTypeId', value = "wins")

# kable(contingencyTable)

DT::renderDataTable({ contingencyTable },rownames= FALSE)

```

**Scatterplot between the loss and home loss with colored by game type, the results show very good linear relationship.**
```{r}
ggplot(franchise_team_total,aes(losses,homeLosses, color=gameTypeId)) +
  geom_point() 
```

**Plot the histogram of the wins by game type**
```{r}
franchise_by_gametype<-franchise_team_total %>% group_by_at(c('teamName','teamId','gameTypeId')) %>% 
    summarise(wins = sum(wins), losses=sum(losses))

ggplot(franchise_by_gametype,aes(gameTypeId,wins)) +
  geom_col() 
```

**Plot the histogram of the wins for each team (side by side with game type)**, *It shows that when the wins in game type=2 is high, the wins in game type=2 is also high.*
```{r}
ggplot(franchise_by_gametype,aes(teamId,wins,gameTypeId)) +
  geom_col() +
  facet_grid(rows=vars(gameTypeId), labeller = label_both,scales = "free",)
```

**The box plot for wins within the game type.**
+ We can see that the median for the game type=2 is around 650 with no outliers
+ The median for the game type=3 is around 70 with two outliers
```{r}

ggplot(franchise_team_total,aes(gameTypeId,wins)) +
  geom_boxplot() +
  facet_grid(rows=vars(gameTypeId), labeller = label_both,scales = "free" )
```

**Create new variable**
+ Over all win percent (winPercent)
+ Win percent at home (homewinPercent)
+ Scatter plot the win percent v.s. home win percent, the result shows strong linear relationship with more outliers than the scatter plot of wins v.s. homewins.
```{r}
#Create new variables of home win percentage and win percentage
franchise_team_percentage<-franchise_team_total %>% replace_na(list(wins= 0, ties = 0, losses=0, homeWins=0, homeLosses=0, homeTies=0)) %>%
  mutate(winPercent=wins/(wins+ties+losses),homewinPercent=homeWins/(homeWins+homeTies+homeLosses)) 

ggplot(data=franchise_team_percentage,aes(winPercent,homewinPercent)) +
  geom_point()
```

