---
title: "Project 1: Vignettes about Reading JSON Files and analysis in R"
author: 
- "Min He"
date: "June 9, 2020 (updated `r Sys.Date()`)"
output: 
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
    code_folding: show
    toc_float:
      collapsed: yes
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE)
library(tidyverse)
library(jsonlite)
library(dplyr)

require(knitr)
library(ggplot2)
```

# JSON Description
JSON (JavaScript Object Notation) is a lightweight data-interchange format. It is easy for humans to read and write. It is easy for machines to parse and generate. JSON is common data formats used to receive data from a web server. JSON is built on two structures: 

+ A collection of name/value pairs.
+ An ordered list of values. 


# Pros/Cons of Packages

We chose the `jsonlite` package to read in our data. The `rjson` package was also considered, but inital attempts needed extra code to implement the data into a dataframe. The `jsonlite` package converts read data directly into a dataframes by default, while `rjson` will convert data into a list by default. Given our need to work with dataframes, `jsonlite` performs better for our needs.


# Data

This data we are using is the NHL records, an jsonlite API is used to retrieve all the team data (including the details about team New Jersey Devils).

## Read the data
```{r echo=TRUE}
# Read in data from JSON source

read_json<-function(root_url,params){
  json_file <- paste0(root_url,params)
  source_data <- fromJSON(json_file)
  franchise_df <- as.data.frame(source_data$data) #Convert to data frame
  return(franchise_df)
}
root_url <- 'https://records.nhl.com/site/api'
franchise_df <- read_json(root_url,'/franchise')
franchise_team_total <- read_json(root_url,'/franchise-team-totals')

franchise_id=23
franchise_team_season <- read_json(root_url,paste0('/franchise-season-records?cayenneExp=franchiseId=',franchise_id))
franchise_team_goalie <- read_json(root_url,paste0('/franchise-goalie-records?cayenneExp=franchiseId=',franchise_id))
franchise_team_skater <- read_json(root_url,paste0('/franchise-skater-records?cayenneExp=franchiseId=',franchise_id))

```

## Basic data exploratory

A basic exploratory data analysis.

+ The scatterplot between losses and home losses shows a very good linear relationship
+ The linear relationship between losses and home losses still hold within each game type.

```{r echo=TRUE}

ggplot(franchise_team_total,aes(losses,homeLosses)) +
  geom_point()
```

```{r echo=TRUE}
ggplot(franchise_team_total,aes(losses,homeLosses)) +
  geom_point() +
  facet_grid(cols=vars(gameTypeId),scales = "free",labeller = label_both)
```

**Take the** *franchiseId=23* **an an example, the scatterplot between most wins in one season and total wins for each team member shows some linear relationship and this relationship is better when the outlier was excluded.**


```{r}
plot(franchise_team_goalie$mostWinsOneSeason, franchise_team_goalie$wins, main='Relationshio between most wins in one season\n v.s. wins for each member', xlab = 'Most win in one season', ylab='Total wins')
plot(franchise_team_goalie$mostWinsOneSeason[-1], franchise_team_goalie$wins[-1], main='Relationshio between most wins in one season\n v.s. wins for each member (exclude outlier)', xlab = 'Most win in one season', ylab='Total wins')
```


## Create Contingency Table
**Show contingency table with columns of team and game type (summarise the total of wins)**
```{r, echo=TRUE}
#Function creates Contingency Table
contingencyTable<-franchise_team_total %>% group_by_at(c('teamName','gameTypeId')) %>% 
    summarise(wins = sum(wins)) %>% 
    spread(key = 'gameTypeId', value = "wins")

kable(contingencyTable)
```

# Plotting

## Scatter plot by color

**Scatterplot between the loss and home loss with colored by game type, the results show very good linear relationship.**

```{r}
ggplot(franchise_team_total,aes(losses,homeLosses, color=gameTypeId)) +
  geom_point() 
```

## Histogram plot by side by side

**Plot the histogram of the wins by game type**

```{r}
franchise_by_gametype<-franchise_team_total %>% group_by_at(c('teamName','teamId','gameTypeId')) %>% 
    summarise(wins = sum(wins), losses=sum(losses))

ggplot(franchise_by_gametype,aes(gameTypeId,wins)) +
  geom_col() 
```


**Plot the histogram of the wins for each team (side by side with game type)**, *It shows that when the wins in game type=2 is high, the wins in game type=2 is also high.*

```{r}
ggplot(franchise_by_gametype,aes(teamId,wins,gameTypeId)) +
  geom_col() +
  facet_grid(rows=vars(gameTypeId), labeller = label_both,scales = "free",)
```

## Box plot side by side

**The box plot for wins within the game type.**

+ We can see that the median for the game type=2 is around 650 with no outliers
+ The median for the game type=3 is around 70 with two outliers

```{r}

ggplot(franchise_team_total,aes(gameTypeId,wins)) +
  geom_boxplot() +
  facet_grid(rows=vars(gameTypeId), labeller = label_both,scales = "free" )
```


# Create new variables

**Create new variable**

+ Over all win percent (winPercent)
+ Win percent at home (homewinPercent)
+ Scatter plot the win percent v.s. home win percent, the result shows strong linear relationship with more outliers than the scatter plot of wins v.s. homewins.


```{r}
#Create new variables of home win percentage and win percentage
franchise_team_percentage<-franchise_team_total %>% replace_na(list(wins= 0, ties = 0, losses=0, homeWins=0, homeLosses=0, homeTies=0)) %>%
  mutate(winPercent=wins/(wins+ties+losses),homewinPercent=homeWins/(homeWins+homeTies+homeLosses)) 

ggplot(data=franchise_team_percentage,aes(winPercent,homewinPercent)) +
  geom_point()
```

# Project Summary

In this project, I read in the NHL data with `JSONLITE` api, and convert the data into dataframe for plotting and exploratory analysis. My initial analysis shows that the wins v.s. the home wins are almost linear related, and this linear relationship holds with game type. 

# Discussion

## what would you do differently?
I would like to transpose the data and create a table contains the result_type ('win','homewin','tie','hometie','loss','homeloss') and another column result_count, and do the analysis for each home v.s. total results.


## what was the most difficult part for you?

I found the most difficult part is to intepreter the win histogram fo each team, as we have 38 teams, it's difficult to see the results when putting them into one graph.

## what are your big take-aways from this project?

My takeways:

+ `jsonlite` is an API easy to use, and the result can be easily converted in dataframe
+ The percentage of win at home is the similar to overall win percentage
+ Game type doesn't have big impact on the win percentage
